Improving cropland maps through tight integration of human and machine intelligence

### 2019-06-21 manuscript/manuscript.Rmd
Paper drafted through imagery section

- name: Lyndon D. Estes
  affilnum: '1'

# - name: Francisco Rodriguez-Sanchez 
#   affilnum: 3
#   email: f.rodriguez.sanc@gmail.com

@@

#   
# - affilnum: 2
#   affil: Big Name University, Department of R, City, BN, 01020, USA
#   
# - affilnum: 3
#   affil: Estacion Biologica de Donana (EBD-CSIC), E-41092 Sevilla, Spain

@@

# Introduction
How much cropland is on the planet, and how much more will be needed to meet humanity's rapidly growing food demands? The first part of this question is unclear, as existing estimates tend to vary widely [e.g. @FritzHighlightingcontinueduncertainty2011; @Fritzneedimprovedmaps2013], which makes it substantially harder to answer the second part. The reason for these difficulties is because cropland estimates rely heavily on remote sensing-derived landcover maps, which can be notoriously high in error, particularly over regions such as Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for an accurate baseline is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015].

@@

We describe here a cropland mapping system that addresses these requirements. The first requirement is enabled by the recent availability of cubesat data that provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017]. Although these data are of lower spectral depth and quality than Landsat, Sentinel, or Worldview imagery, they enable country-continent scale image mosaics to be created for multiple periods during the crop growing calendars, and this representation of intra-annual variability is more important for cropland classification than spectral depth [@Debatsgeneralizedcomputervision2016]. Additionally, despite this imagery being up to ~16 times coarser than much of Bing or Google imagery, it is fine enough for humans to be able to discern fields with sufficient confidence [e.g. see @Esteslargeareaspatiallycontinuous2018; @FourieBetterCropEstimates2009]. This allows labels to be generated on the same imagery processed by the classifier, thereby addressing one of the two needs related to training data (requirement 3). The second requirement is addressed by a computer vision/machine learning classifer that is effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016] and was re-engineered to run on high performance, cloud-based computing clusters. This classifier is tightly coupled to a front-end platform for collecting label data, which includes rigorous accuracy assessment protocols [@Estesplatformcrowdsourcingcreation2016] and a novel approach for merging multiple maps into consensus labels, thereby minimizing image interpretation error. The training and machine learning components are combined within an "active learning" framework, wherein the machine learning process assesses classification uncertainty in unlabelled area after a training step, and selects sites from areas of highest uncertainty for additional labelling [@DebatsIntegratingactivelearning2017]. Our framework automates this interactive process, which has been shown to be important for improving accuracy in previous studies [e.g. @XiongNominal30mCropland2017; others add]. Furthermore, it is more efficient than purely randomized or conventionally stratified approaches to label collection, meaning that fewer labels are needed to generate a given level of accuracy [@DebatsIntegratingactivelearning2017].

We demonstrate this approach to map cropland in Ghana, a country where agriculture is extensive and dominated by smallholder production.

@@

# Methods

- Imagery and image processing methods
- Sample grid
- The system
    - Crowdsourcing platform
          - Training and validation data
          - Accuracy assessment
    - Machine learning
          - Feature engineering
          - ML pipeline Apache Spark
    - Iteration cycle
    - Segmentation

>>>>

- name: \*Provisional author list\*
  affilnum: '1'

- name: Lyndon D. Estes
  affilnum: '1'
  email: lestes@clarku.edu

- name: Su Ye
  affilnum: '1'

- name: Lei Song
  affilnum: '1'

- name: Ron Eastman
  affilnum: '1'

- name: Sitian Xiong
  affilnum: '1'

- name: Tammy Woodard
  affilnum: '1'

- name: Boka Luo
  affilnum: '1'

- name: Dennis McRitchie
  affilnum: '2'

- name: Ryan Avery
  affilnum: '3'

- name: Kelly Caylor
  affilnum: '3'

- name: Stephanie Debats
  affilnum: '4'

- name: SpatialCollective
  affilnum: '5'

- name: Meridia
  affilnum: '6'

- name: Azavea
  affilnum: '7'

@@

- affilnum: 2
  affil: Dennis and Sons, Tucson, AZ, USA

- affilnum: 3
  affil: UCSB

- affilnum: 4
  affil: Uber

- affilnum: 5
  affil: SpatialCollective

- affilnum: 5
  affil: Meridia

- affilnum: 5
  affil: Azavea

@@

# Introduction
How much cropland is on the planet, and where is it located? How much more will be needed to meet humanity's rapidly growing food demands? The answer to the first question is unclear, as existing estimates tend to vary widely and cropland maps show substantial spatial disagreements [e.g. @FritzHighlightingcontinueduncertainty2011; @Fritzneedimprovedmaps2013], which makes it difficult to answer the second. The first question can't be clearly answered because cropland estimates rely heavily on remote sensing-derived landcover maps, which can be notoriously high in error, particularly over regions such as Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for an accurate baseline is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015]. *[Fill in more reason why cropland maps are important]*

@@

We describe here a cropland mapping system that addresses these requirements on the way to creating an object-based map of individual fields The first requirement is enabled by the recent availability of cubesat data that provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017]. Although these data are of lower spectral depth and quality than Landsat, Sentinel, or Worldview imagery, they enable country-continent scale image mosaics to be created for multiple periods during the crop growing calendars, and this representation of intra-annual variability is more important for cropland classification than spectral depth [@Debatsgeneralizedcomputervision2016]. Additionally, despite this imagery being up to ~16 times coarser than much of Bing or Google imagery, it is sufficiently resolved for humans to discern most fields [e.g. see @Esteslargeareaspatiallycontinuous2018; @FourieBetterCropEstimates2009]. This allows labels to be generated on the same imagery processed by the classifier, thereby addressing one of the two needs related to training data (requirement 3). The second requirement is addressed by a computer vision/machine learning classifer that is effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016] and was re-engineered to run on high performance, cloud-based computing clusters. This classifier is tightly coupled to a front-end platform for collecting label data, which includes rigorous accuracy assessment protocols [@Estesplatformcrowdsourcingcreation2016] and a novel approach for merging multiple maps into consensus labels, thereby minimizing image interpretation error. The training and machine learning components are combined within an "active learning" framework, wherein the machine learning process assesses classification uncertainty in unlabelled areas after a training step, and selects sites from areas of highest uncertainty for additional labelling [@DebatsIntegratingactivelearning2017]. Our framework automates and formalizes this interactive approach to label selection, which has been shown to be important for improving accuracy in previous studies [e.g. @XiongNominal30mCropland2017; others add]. Furthermore, it is more efficient than purely randomized or conventionally stratified approaches to label collection, meaning that fewer labels are needed to generate a given level of accuracy [@DebatsIntegratingactivelearning2017]. Finally, an unsupervised  segmentation step is applied to the imagery and merged with the pixel-wise classifications from the machine learning process, **resulting in a further sharpening of accuracy** and resulting in a final vector-based field boundary map.

We demonstrate this approach to map cropland in Ghana, a country where agriculture is predominantly small-scale and comprises a broad mix of agricultural systems, including large areas where shifting agricultural practices predominate [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016]. These factors make Ghana emblematic of the substantial challenges facing cropland classification over the broader region.

@@

# Methods
## System overview 
The mapping system is centered around four major components: 1) an image acquisition and pre-processing component, 2) a training data collection platform, 3) a machine learning component, and a 4) segmentation component. The first component is applied as a once-off step to generate an image catalog necessary for the mapping geography, while the second component initiates the the interactive image training and classification (component 3) pipeline that results in a pixel-wise posterior probability map of cropland probabilities on its final iteration (Fig. 1). The final segmentation step (component 4) is then initiated and applied to both the pre-processed primary images and the posterior probability maps, resulting in the final vectorized fields boundary maps.

## System components
### Imagery
The primary image source used in our system is PlanetScope Analytic surface reflectance product [@PlanetTeamPlanetApplicationProgram2017], although can be readily adapted to apply to Sentinel and Landsat images. PlanetScope provides three visual (red, green, blue) and a near-infrared band at 3-4 m resolution at nominal daily frequency. Although these images are substantially pre-preprocessed and corrected for atmospheric effects, there are substantial residual errors from inter-sensor differences and the radiometric normalization process [@HouborgDailyRetrievalNDVI2018], substantial variation in the orientation of scene footprints, as well as a high frequency of cloud cover over the study region [@WilsonRemotelySensedHighResolution2016]. To correct for these factors, we developed a procedure for creating temporal composites representing the primary growing and non-growing seasons within a single year.

```{r, echo = FALSE, out.width="50%", fig.cap="An overview of the primary system components, the data stores that hold the inputs and outputs from each component, and the direction of connections between them. The dashed line indicates iterative interactions, while solid lines indicate one-time or irregular connections.", fig.align='center'}
knitr::include_graphics('figures/figure1.png')
```

PlanetScope imagery is accessed via the Planet API, and an initial order is placed wherein the imagery representing all available dates in the two compositing periods and within a bounding box covering the mapping geography is collected and transferred directly to the system's cloud storage (hosted on Amazon Web Services S3). Imagery that don't having certain quality flags are excluded (e.g. failing to meet minimum spatial precision).

We then transform all images into an "analysis ready data" (ARD) product (cite) that is prepared against a reference grid that is situated within a larger 1 degree resolution grid that covers the entire continent. Each 1 degree cell intersecting the mapping geography (Fig. 2A) is divided into 400 0.05 degree cells (Fig. 2B), which are the reference cells for image compositing. The ARD process then clips the imagery intersecting each 0.05 cell. 
<!-- and then applies the usable data mask (UDM) supplied with the imagery to remove low quality and cloudy pixels.  -->

```{r, echo = FALSE, out.width="100%", fig.cap="Figure scales", fig.align='center'}
knitr::include_graphics('figures/figure2.png')
```
The temporal compositing process is then applied to the ARD imagery, and is done for two time periods. In Ghana, a composite is made to cover the dry season period (December-February) and to span the entire wet season (May - September).  The longer period is necessary for the wet season because of the high frequency of cloud cover, which severely limits the number of clear scenes over any given area. For each image in the compositing process, we pass a 3X3 median filter over the image to remove speckling, and then calculate a weight value based on a simplified formula of the Haze Optimized Transformation (HOT) index [@Zhangimagetransformcharacterize2002; @ZhuObjectbasedcloudcloud2012]:

$$
HOT_{weight} = \frac{1}{(blue - 0.5red)^2}
$$
This assigns a low weight to pixels that are cloudy, hazy, or smoky pixels on a particular date. We then use the weights to calculate a weight mean for each pixel and each band for the final composite for each time period. The composites are then added to the S3 store (Fig. 1), and a "slippy map" rendering is created for each composite using RasterFoundry (cite). This web-based rendering of the image is used in the training data platform (see section X).

### Training data platform
- Training and validation data
- Accuracy assessment

### Classification pipeline
- Feature engineering
- ML pipeline Apache Spark

### Segmentation

## The active learning loop
